When are Fuzzy Extractors Possible?

### 2014-10-06 trunk/info-bounds/FuzzyExtPossibility.tex
edited the low-dim discussion in 1.1 and sufficiency in 1.2

\begin{table}[h]
\begin{center}
\begin{tabular}{c  l l }
 & Known Distribution & Family of Distributions\\
\hline
Secure Sketch & Yes~(\thref{cons:leveling}) & No~(\thref{thm:imposs sketch})\\
\hline
Fuzzy Extractor & Yes~(\corref{cor:extension to fuzz ext}) & No~(\thref{thm:imposs fuzz ext}) 
\end{tabular}
\end{center}
\caption{Is fuzzy min-entropy sufficient to extract a superlogarithmic length key?  Results are information-theoretic.}
\label{tab:main results}
\end{table}

@@

\paragraph{Techniques for Negative Results for Distribution Families}
We construct a family of distributions $\mathcal{W}$ and prove impossibility for a uniformly random $W\in \mathcal{W}$~(instead of proving impossibility for a worst-case $W$).
We start by observing the following asymmetry: $\gen$  sees only the sample $w$ (obtained via $W\leftarrow \mathcal{W}$ and $w\leftarrow W$), while
the adversary knows $W$.   To exploit the asymmetry, we construct $\mathcal{W}$ so that conditioning on the knowledge of $W$ reduces the distribution to a single affine line, but conditioning on $w$ leaves the rest of the distribution uniform on the entire space. \bnote{this isn't quite accurate, the first coordinate is known.  you have no info on the rest of the support}

@@

The result for fuzzy extractors is delicate.  The fact that $p$ partitions the space~(into non overlapping regions) relies on perfect correctness.  Extending this result to imperfect correctness seems challenging and is an interesting open problem.
The result also relies on the majority of points lying near boundaries of a set. Indeed, fuzzy extractors for arbitrary distributions are known in metric spaces where the boundary of a set is not the majority of its area.\bnote{cite cite cite!!!}

\subsection{Sufficiency of $\Hfuzz(W)$ in related settings}
\label{sec:related settings}
\bnote{this needs work}
We have evidence that fuzzy min-entropy is sufficient in settings related to information-theoretic fuzzy extractors.  
%We first consider the interactive setting.  
First, an alternate definition fuzzy extractors and secure sketches may provide computational security instead of information-theoretical security~\cite{fuller2013computational}.  In this setting, under the strong assumption of semantically-secure multilinear maps, fuzzy extractors and secure sketches can be constructed for any distribution $W$~\cite{BitanskyCKP14}.  We note that the secure sketch of~\cite{BitanskyCKP14} provides computational unpredictability of $W |p$.  Our negative result that implies that $W|p$ cannot have pseudoentropy~(discussion in \secref{sec:feas comp sec sketch}).

Second, we can weaken the computational assumption necessary for $\Hfuzz(W)$ to be sufficient by moving to an interactive setting.
Consider a user, Alice, using a noisy secret to authenticate with a remote server, Bob.  In this setting, Alice could initially send the value $w$ to Bob and then authenticate with subsequent value $w'$.  If the server holds $w$ and the user $w'$ the two parties can agree on a key using an interactive protocol.
%\footnote{It is dangerous to allow a remote server access to original reading $w$ as they can then impersonate the user.  Ideally, the server would keep a ``hashed'' value of $w$.  This is  done for noiseless secrets like passwords.}  
In this world, secure two-party computation can produce a key whenever the distribution $W$ has fuzzy min-entropy.  The minimal assumption necessary for two party computation is oblivious transfer~\cite{yao1986generate,lindell2009proof}.\lnote{OT, right?; we can ask Ran about their paper with Tal R. or read the paper}

@@

\noindent
In our positive results we considered an arbitrary finite metric space.  The relevant parameters were the size of balls and the size of the metric space.  Our negative results are specific to the Hamming metric.
\section{Insufficiency of $\Hfuzz(W)$ to secure sketch from distribution families}
\label{sec:family of dist}

@@

\begin{itemize}
\item \propref{prop:each element good}: for all $W\in \mathcal{W}$, $\Hfuzz(W) = \omega (\log n)$. That is, $\forall z, \Hfuzz(V | Z=z) = \omega(\log n)$.
\item \propref{prop:distribution uniform}: the distribution $V$ is uniform.
\item \lemref{lem:secure sketch entropy loss}: for any secure sketch on $V$, the support size of $V | \sketch(V)$ decreases significantly.  Here we show the minimum distance of $V|\sketch(V)$ is at least $t$.
\item \lemref{lem:side info determines sketch}:  for most lines $Z$, the intersection of the support of $V|Z$ and $V | \sketch(V)$ is small.  \\That is, $\tilde{H}_0(V | \sketch(V), Z) < 2$.
\end{itemize}
\end{proof}
\noindent
\textbf{Note:} There is tradeoff between the size of $\mathbb{F}$ and the error tolerance required for the counter example.  By increasing $t$ it is possible to show a counter example for a smaller $\mathbb{F}$.
\subsubsection{Implications for Computational secure sketches}
\label{sec:feas comp sec sketch}
Fuller et al. showed that computational secure sketches that provide pseudoentropy imply information-theoretic secure sketches with almost the same parameters~\cite[Corollary 3.8]{fuller2013computational}.  %This means that ruling out any information-theoretic secure sketches also rules out a computational secure sketch.  
The definitional of Fuller et al. uses a weak version of pseudoentropy~\cite{DBLP:journals/siamcomp/HastadILL99} due to Gentry and Wichs~\cite{gentry2011separating}.

@@

A \emph{HILL secure sketch} is obtained by replacing the security condition in \defref{def:secure sketch} with HILL entropy~(adding parameters $\epsilon_{sec}, s_{sec}$ parameters to the definition, representing the distinguishing circuit).  By the contrapositive of~\cite[Corollary 3.8]{fuller2013computational}, no sketch can retain HILL entropy for the same family of distributions:

@@

\section{Insufficiency of $\Hfuzz(W)$ to fuzzy extract from distribution families}
\label{sec:imposs fuzz ext}
In the previous section, we showed a family of distributions that does not admit a secure sketch.  We provide a similar result for fuzzy extractors.

@@

The optimum strategy is to include these values uniformly from different $Z$ values.
We show this across different sketch values.  Consider some fixed sketch value $s$ and let $h_s= H_0(V | \sketch(V) = s)$.  %That is, there are $2^{h_s}$ possible values for $V$ conditioned on $\sketch(V) = s$.  
Recall that 
\[
\tilde{H}_0(V | \sketch(V)) =  \log \expe_{s\in \sketch(V)} 2^{H_0(V | \sketch(V) = s)}  = \log \expe_{s\in \sketch(V)} 2^{h_s} %\le   (\gamma-t+1)\log |\mathbb{F}|+1.
\]  
Conditioned on seeing the point $V$ there are $(|\mathbb{F}|-1)^{\gamma-1}/|\mathbb{F}|$ possible values for $Z$ with disjoint support outside of the sketched point.  Consider these possible values for $Z$ as bins to be filled with the $2^{h_{ss}}$ balls~(possible values of $V | \sketch(V)=ss$).  Each bin receives automatically receives one free point~(all the distributions share $v$).  The average number of balls in each bin is maximized when the bins are filled equally.  That is, the average number of balls in each bin is bounded by the number of balls divided by the number of bins.  That is, 
\begin{align*}
\tilde{H}_0(V |Z  , \sketch(V) = ss) &\le \log \left(\frac{\text{\# balls}+\text{\# bins}}{\text{\# bins}}\right)\\
&= \log \left(\frac{2^{h_{ss}}|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} +1 \right)
\end{align*}
Then averaging over the possible values of $s$, we have the following as long as $t\ge 4$~(using  \lemref{lem:log-minus-one}, which appears below):
\begin{align*}
\tilde{H}_0(V |Z , \sketch(V) ) &= \log \expe_{s\in \sketch(V)} 2^{\tilde{H}_0(V |  \sketch(V) =ss , (Z| \sketch(V) =ss) )}\\
&= \log\expe_{s\in \sketch(V)} \left(\frac{2^{h_s}|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} +1\right)\\
&\le \max\left\{ \log \left(\frac{|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \expe_{s\in \sketch(V)} 2^{h_s}\right)+1, 1\right\}.
\end{align*}
Where the inequality follows because $\log x+1 \le \max\{ 1+ \log x,1\}$ for $x\ge 0$.
The left operand to $\max$ is bounded by $2$~(bounded the $\max$ by $2$):
\begin{align*}
\log \left(\frac{|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \expe_{s\in \sketch(V)} 2^{h_s}\right)+1
&=\log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + \log \left(\expe_{s\in \sketch(V)} 2^{h_s}\right) +1\\
&=\log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + \tilde{H}_0(V | \sketch(V)) +1 \\ 
&\le \log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + (\gamma-t+1)\log |\mathbb{F}|+2\\
&\le (\gamma-t+2)\log |\mathbb{F}| - (\gamma-1) \log (|\mathbb{F}|-1)+2\\
&< (\gamma-t+2)\log |\mathbb{F}| - (\gamma-2) \log |\mathbb{F}| +2 \ \ \ \ \mbox{(by \lemref{lem:log-minus-one})}\\
&\le (4-t)\log |\mathbb{F}| +2< 2\,.
\end{align*}
\end{proof}

@@

This means that the $\gen$ algorithm may include points for possible $Z$ values into parts $Q_{p^*, \key^*}$~(other than $v$) and these values are disjoint.  The optimum strategy is to include these values uniformly from different $Z$ values.  Consider the set of all preimages of $R_{small}$ denoted $Q_{small} = \cup_{\key\in R_{small}} \inter(Q_{\key, p^*})$.  Note that $Q_{small} \le 2^{n-4\nu}|R_{small}|$.  We now show that the intersection between $Q_{\key, p^*}$ is small for most possible values $z$.  As before each bin~(the values of $z$)  receives one ball for free~(the point $v$).
\begin{align*}
\expe_z |Q_{small} \cap (V | P=p^* \wedge Z=z) | &\le \left(\frac{\text{\# balls}+\text{\# bins}}{\text{\# bins}}\right)\\
&\le \frac{2^{n-4\nu}|R_{small}|}{2^{n - 3\nu}}+1\\
&=\frac{|R_{small}|}{2^{\nu}}+1
\end{align*}
In expectation across $Z$, 
\[\frac{\frac{|R_{small}|}{2^{\nu}}+1}{|R_{small}|} \le \frac{1}{2^\nu}+\frac{1}{|R_{small}|} \] fraction of $R_{small}$ receive any support.  %Thus, at most $1+1/2^{\nu} = \ngl(n)$ keys in $R_{small}$ have any support  conditioned on $p^*$ and $Z$~(note this is an expectation across the values of $Z$).  
We now present a distinguisher $D_{p^*}$ for a particular $p^*$:
\begin{enumerate}
\item On input $x, z$.
%\item If $x\not \in R_{small}$ output random bit $b$.
\item Compute $V|P=p^* \wedge Z=z$ and $Q_{p^*, x}$. 
\item If $(Q_{p^*, x} \cap V|P=p^* \wedge Z=z) =\emptyset$ output $b=0$.
\item Else output $b=1$.
\end{enumerate}

>>>>

\begin{table}[h]
\begin{center}
\begin{tabular}{c  l l }
 & Known Distribution & Family of Distributions\\
\hline
Secure Sketch & Yes~(\corref{cor:extension to fuzz ext}) & No~(\thref{thm:imposs sketch})\\
\hline
Fuzzy Extractor & Yes~(\corref{cor:extension to fuzz ext}) & No~(\thref{thm:imposs fuzz ext}) 
\end{tabular}
\end{center}
\caption{Is fuzzy min-entropy sufficient to extract a superlogarithmic length key?  Results are information-theoretic.}
\label{tab:main results}
\end{table}

@@

\paragraph{Techniques for Negative Results for Distribution Families}
We construct a family of distributions $\mathcal{W}$ and prove impossibility for a uniformly random $W\in \mathcal{W}$~(instead of proving impossibility for a worst-case $W$).
We start by observing the following asymmetry: $\gen$  sees only the sample $w$ (obtained via $W\leftarrow \mathcal{W}$ and $w\leftarrow W$), while
the adversary knows $W$.   To exploit the asymmetry, we construct $\mathcal{W}$ so that conditioning on the knowledge of $W$ reduces the distribution to a single affine line, but conditioning on $w$ leaves the rest of the distribution uniform on a large fraction of the entire space.

@@

The result for fuzzy extractors is delicate.  It uses the fact that $p$ partitions the space into nonoverlapping regions, which is implied by perfect correctness.  Extending this result to imperfect correctness seems challenging and is an interesting open problem. It also uses the fact that there are few points far from the boundary of every region, which is implied by the geometry of  the high-dimensional Hamming space.  This fact seems crucial: in contrast, in low-dimensional Euclidean space, which does not have this property, a single fuzzy extractor can work for any distribution with sufficient $\Hfuzz$. (Such a construction would use quantization or tiling, similar to, for example, \cite{CK03,LT03,CZC04,LC06,BDHTV10,VTOSS10}.  Each sample from $W$ would map to the ``tile'' containing it, from which the output key would be extracted. A randomly chosen quantizer could be used to ensure that few samples lie near the boundary, giving almost-perfect correctness; if perfect correctness is desired, we can give up on security for those rare samples and simply use a special value of $p$ to indicate that one of them was the input.)

\subsection{Sufficiency of $\Hfuzz(W)$ in related settings}
\label{sec:related settings}
The security definition of fuzzy extractors and secure sketches can be weakened to require protection only against computationally bounded adversaries~\cite{fuller2013computational}.   In this computational setting, fuzzy extractors and secure sketches can be constructed for the family of all distributions $W$ with superlogarithmic $\Hfuzz$ by using virtual grey-box obfuscation for all circuits~\cite{BitanskyCKP14}. The construction places into the $p$ the obfuscated program for testing proximity to $w$ and outputting the appropriate value if the test passes.  In addition to relying on strong assumptions for security (namely, the existence of semantically-secure multilinear maps), this construction is not of practical efficiency. Note that if this construction is used for a secure sketch, $W$ will remain unpredictable conditioned on $p$, but will not have pseudoentropy (see  \secref{sec:feas comp sec sketch} for details).

Furthermore, the functional definition of fuzzy extractors and secure sketches can be weakened to permit interaction between the party having $w$ and the party having $w'$. Such a weakening is useful for secure remote authentication~\cite{Boyen05secureremote}. When both interaction and computational assumptions are allowed, secure two-party computation can produce a key that will be secure whenever the distribution $W$ has fuzzy min-entropy.  The two-party computation protocol needs to be secure without assuming authenticated channels; it can be built under the assumptions that collision-resistant hash functions and enhanced trapdoor permutations exist~\cite{DBLP:journals/joc/BarakCLPR11}.

@@

\noindent
In our positive results we considered an arbitrary finite metric space.  The relevant parameters were the size of balls and the size of the metric space.  Our negative results are specific to the Hamming metric.
\section{Impossibility of secure sketches for all families with $\Hfuzz$}
\label{sec:family of dist}

@@

\begin{itemize}
\item \propref{prop:each element good}: for all $W\in \mathcal{W}$, $\Hfuzz(W) = \omega (\log n)$. That is, $\forall z, \Hfuzz(V | Z=z) = \omega(\log n)$.
\item \propref{prop:distribution uniform}: the distribution $V$ is uniform.
\item \lemref{lem:secure sketch entropy loss}: for any secure sketch on $V$, the support size of $V | \sketch(V)$ decreases significantly.  Here we show the minimum distance of $V|\sketch(V)$ is at least $t$.
\item \lemref{lem:side info determines sketch}:  for most lines $Z$, the intersection of the support of $V|Z$ and $V | \sketch(V)$ is small.  \\That is, $\tilde{H}_0(V | \sketch(V), Z) < 2$.
\end{itemize}
\end{proof}
\noindent
\textbf{Note:} There is tradeoff between the size of $\mathbb{F}$ and the error tolerance required for the counter example.  By increasing $t$ it is possible to show a counter example for a smaller $\mathbb{F}$.
\subsection{Implications for Computational secure sketches}
\label{sec:feas comp sec sketch}
Fuller et al. showed that computational secure sketches that provide pseudoentropy imply information-theoretic secure sketches with almost the same parameters~\cite[Corollary 3.8]{fuller2013computational}.  %This means that ruling out any information-theoretic secure sketches also rules out a computational secure sketch.  
The definition of Fuller et al. uses a weak version of pseudoentropy~\cite{DBLP:journals/siamcomp/HastadILL99} due to Gentry and Wichs~\cite{gentry2011separating}.

@@

A \emph{HILL secure sketch} is obtained by replacing the security condition in \defref{def:secure sketch} with HILL entropy~(adding parameters $\epsilon_{sec}, s_{sec}$ to the definition, representing the distinguishing circuit).  By the contrapositive of~\cite[Corollary 3.8]{fuller2013computational}, no sketch can retain HILL entropy for the same family of distributions:

@@

\section{Impossibility of fuzzy extractors for all families with $\Hfuzz$}
\label{sec:imposs fuzz ext}
In the previous section, we showed a family of distributions that does not admit a secure sketch.  We provide a similar result for fuzzy extractors.

@@

The optimum strategy is to include these values uniformly from different $Z$ values.
We show this across different sketch values.  Consider some fixed sketch value $s$ and let $h_s= H_0(V | \sketch(V) = s)$.  %That is, there are $2^{h_s}$ possible values for $V$ conditioned on $\sketch(V) = s$.  
Recall that 
\[
\tilde{H}_0(V | \sketch(V)) =  \log \expe_{s\in \sketch(V)} 2^{H_0(V | \sketch(V) = s)}  = \log \expe_{s\in \sketch(V)} 2^{h_s} %\le   (\gamma-t+1)\log |\mathbb{F}|+1.
\]  
Conditioned on seeing the point $V$ there are $(|\mathbb{F}|-1)^{\gamma-1}/|\mathbb{F}|$ possible values for $Z$ with disjoint support outside of the sketched point.  Consider these possible values for $Z$ as containers to be filled with the $2^{h_{ss}}$ items~(possible values of $V | \sketch(V)=ss$).  Each container receives automatically receives one free point~(all the distributions share $v$).  The average number of items in each container is maximized when the containers are filled equally.  That is, the average number of items in each container is bounded by the number of items divided by the number of container.  That is, 
\begin{align*}
\tilde{H}_0(V |Z  , \sketch(V) = ss) &\le \log \left(\frac{\text{\# items}+\text{\# containers}}{\text{\# containers}}\right)\\
&= \log \left(\frac{2^{h_{ss}}|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} +1 \right)
\end{align*}
Then averaging over the possible values of $s$, we have the following as long as $t\ge 4$~(using  \lemref{lem:log-minus-one}, which appears below):
\begin{align*}
\tilde{H}_0(V |Z , \sketch(V) ) &= \log \expe_{s\in \sketch(V)} 2^{\tilde{H}_0(V |  \sketch(V) =ss , (Z| \sketch(V) =ss) )}\\
&= \log\expe_{s\in \sketch(V)} \left(\frac{2^{h_s}|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} +1\right)\\
&\le \max\left\{ \log \left(\frac{|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \expe_{s\in \sketch(V)} 2^{h_s}\right)+1, 1\right\}.
\end{align*}
Where the inequality follows because $\log x+1 \le \max\{ 1+ \log x,1\}$ for $x\ge 0$.
The left operand to $\max$ is bounded by $2$~(bounded the $\max$ by $2$):
\begin{align*}
\log \left(\frac{|\mathbb{F}|}{(|\mathbb{F}|-1)^{\gamma-1}} \expe_{s\in \sketch(V)} 2^{h_s}\right)+1
&=\log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + \log \left(\expe_{s\in \sketch(V)} 2^{h_s}\right) +1\\
&=\log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + \tilde{H}_0(V | \sketch(V)) +1 \\ 
&\le \log |\mathbb{F}| - (\gamma -1)\log (|\mathbb{F}|-1) + (\gamma-t+1)\log |\mathbb{F}|+2\\
&\le (\gamma-t+2)\log |\mathbb{F}| - (\gamma-1) \log (|\mathbb{F}|-1)+2\\
&< (\gamma-t+2)\log |\mathbb{F}| - (\gamma-2) \log |\mathbb{F}| +2 \ \ \ \ \mbox{(by \lemref{lem:log-minus-one})}\\
&\le (4-t)\log |\mathbb{F}| +2< 2\,.
\end{align*}
\end{proof}

@@

This means that the $\gen$ algorithm may include points for possible $Z$ values into parts $Q_{p^*, \key^*}$~(other than $v$) and these values are disjoint.  The optimum strategy is to include these values uniformly from different $Z$ values.  Consider the set of all preimages of $R_{small}$ denoted $Q_{small} = \cup_{\key\in R_{small}} \inter(Q_{\key, p^*})$.  Note that $Q_{small} \le 2^{n-4\nu}|R_{small}|$.  We now show that the intersection between $Q_{\key, p^*}$ is small for most possible values $z$.  As before each container~(the values of $z$)  receives one item for free~(the point $v$).
\begin{align*}
\expe_z |Q_{small} \cap (V | P=p^* \wedge Z=z) | &\le \left(\frac{\text{\# items}+\text{\# containers}}{\text{\# containers}}\right)\\
&\le \frac{2^{n-4\nu}|R_{small}|}{2^{n - 3\nu}}+1\\
&=\frac{|R_{small}|}{2^{\nu}}+1
\end{align*}
In expectation across $Z$, 
\[\frac{\frac{|R_{small}|}{2^{\nu}}+1}{|R_{small}|} \le \frac{1}{2^\nu}+\frac{1}{|R_{small}|} \] fraction of $R_{small}$ receive any support.  %Thus, at most $1+1/2^{\nu} = \ngl(n)$ keys in $R_{small}$ have any support  conditioned on $p^*$ and $Z$~(note this is an expectation across the values of $Z$).  
We now present a distinguisher $D_{p^*}$ for a particular $p^*$:
\begin{enumerate}
\item On input $x, z$.
%\item If $x\not \in R_{small}$ output random bit $b$.
\item Compute $V|P=p^* \wedge Z=z$ and $Q_{p^*, x}$. 
\item If $(Q_{p^*, x} \cap V|P=p^* \wedge Z=z) =\emptyset$ output $b=0$.
\item Else output $b=1$.
\end{enumerate}
